---
title: "Cysteine S-acetylation is a post-translational modification involved in metabolic
  regulation"
author: "Akshay Bareja"
date: "2025-06-30"
output: html_document
output: 
  html_document:
    code_folding: hide
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(quantro)
library(pheatmap)
library(ggrepel)
library(FactoMineR)
library(factoextra)
library(limma)
library(org.Mm.eg.db)
library(fgsea)
library(VennDiagram)
```

### Helper function

```{r message = FALSE, warning = FALSE}
heatmap_function <- function(df, 
                         cols, 
                         log_transform,
                         cluster_number,
                         distance_metric_row,
                         distance_metric_col,
                         cluster_cols = TRUE,
                         agglomeration_method,
                         sample){
  df_select <- df %>% 
    dplyr::select(unique_id, all_of(cols)) %>% 
    tidyr::drop_na() # remove all NAs
  
  highest_variance <- df_select %>% 
    dplyr::rowwise() %>% 
    dplyr::mutate(variance = var(c_across(starts_with("Abundance")))) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(desc(variance)) %>% 
    dplyr::slice(1L) %>% 
    dplyr::pull(unique_id)
  
  lowest_variance <- df_select %>% 
    dplyr::rowwise() %>% 
    dplyr::mutate(variance = var(c_across(starts_with("Abundance")))) %>% 
    dplyr::ungroup() %>% 
    dplyr::arrange(variance) %>% 
    dplyr::slice(1L) %>% 
    dplyr::pull(unique_id)
  
  # convert to matrix
  df_mat <- as.matrix(df_select[, -1])
  
  # optional log transform
  if(log_transform){
    df_mat <- apply(df_mat, 2, log2)
  }
  
  # scale data
  df_scaled <- t(scale(t(df_mat))) %>% 
                  data.frame()
  df_mat_scaled <- df_scaled %>% 
                    tidyr::drop_na() %>% # because the dist() function below does not tolerate NAs or NaN. drop_na() gets rid of both!
                    as.matrix()

  # create column annotations
  my_sample_col <- data.frame(
    sample
  )
  
  # add sample names as rownames
  rownames(my_sample_col) <- colnames(df_mat_scaled)
  
  pheatmap(df_mat_scaled,
           annotation_col = my_sample_col,
           cutree_rows = cluster_number,
           clustering_distance_rows = distance_metric_row,
           clustering_distance_cols = distance_metric_col,
           clustering_method = agglomeration_method,
           cluster_cols = cluster_cols,
           show_rownames = FALSE,
           show_colnames = FALSE, 
           main = paste0("Splitting Analytes into ", cluster_number," clusters"))
  
  my_results <- list("df_mat_scaled" = df_mat_scaled,
                     "highest_variance" = highest_variance, 
                     "lowest_variance" = lowest_variance)
  #return(my_results) 
}

# Example use
# a <- heatmap_function(df = all_acetyl_cys_peptide, cols = c(24:29), log_transform = FALSE, cluster_number = 6, distance_metric_row = "euclidean",distance_metric_col = "euclidean", agglomeration_method = "complete", sample = c("Liver", "Lung", "Kidney", "Skm", "Fr_liver", "Fr_BAT"))
```

## Figure 1

Incubation with Acetyl-CoA increases cysteine acetylation. 

### Import data

```{r message = FALSE, warning = FALSE}
liver_acetyl_peptide <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20221220EKK20220908_Liver+AcetylCoA_PD20220112.xlsx")

# Add a unique identifier to each row
liver_acetyl_peptide$peptide_id <- paste0("analyte_", 1:nrow(liver_acetyl_peptide))

# import protein-level data
liver_acetyl_protein <- read_excel("Protein data/All Proteins _ all columns _ 20221220EKK20220908_Liver+AcetylCoA_PD20220112.xlsx")

# In the case of peptide-level data, there are some peptides that match to MULTIPLE proteins
# Decision - treat these as DIFFERENT peptides

# select relevant columns 
liver_acetyl_peptide <- liver_acetyl_peptide %>% 
                            dplyr::select(peptide_id, `Master Protein Accessions`, starts_with("Abundances (Normalized)")) %>% 
                            tidyr::drop_na()

# reshape data frame
liver_acetyl_peptide_long <- liver_acetyl_peptide %>%
  tidyr::separate_longer_delim(`Master Protein Accessions`,
                               delim = ";") %>%
  tidyr::pivot_longer(
    cols = starts_with("Abundance"),
    names_to = "sample_name",
    values_to = "peptide_value"
  ) %>%
  dplyr::rename(Accession = `Master Protein Accessions`) %>%
  dplyr::mutate(Accession = str_trim(Accession))

# select relevant columns
liver_acetyl_protein <- liver_acetyl_protein %>% 
                            dplyr::select(Accession, starts_with("Abundances (Normalized)")) %>% 
                            tidyr::drop_na()

# reshape data frame
liver_acetyl_protein_long <- liver_acetyl_protein %>% 
  tidyr::pivot_longer(cols = starts_with("Abundance"),
                      names_to = "sample_name",
                      values_to = "protein_value") 

# join data frames, normalize peptide value, and reshape
liver_acetyl_peptide_norm <- liver_acetyl_peptide_long %>% 
  dplyr::inner_join(liver_acetyl_protein_long,
                    by = c("Accession", "sample_name")) %>%
  dplyr::mutate(normalized_value = peptide_value/protein_value,
                   .keep = "unused") %>% 
  dplyr::distinct(peptide_id, Accession, sample_name, 
                  .keep_all = TRUE) %>% # NOTE: if you don't do this, you end up with list columns because of non-unique rows!!
  # This is likely happening because of too many many-to-many matches
  tidyr::pivot_wider(
    names_from = sample_name,
    values_from = normalized_value
  ) %>% 
  tidyr::unite("unique_id", peptide_id:Accession) %>%  # this way every entry will be unique
  dplyr::select(unique_id, contains("TCEP"))
# Note that these are a few analytes that have 1s across all samples, indicating that in these cases the peptide and protein abundance are IDENTICAL
 
# create a column called `zero_var` that identifies those rows with 0 variance i.e. all 1s
liver_acetyl_peptide_norm <- liver_acetyl_peptide_norm %>% 
                                  dplyr::mutate(zero_var = if_all(starts_with("Abundances"), ~ . == 1))

liver_acetyl_peptide_norm_var <- liver_acetyl_peptide_norm %>% 
                                    dplyr::filter(!zero_var) %>% 
                                    dplyr::select(-zero_var)
liver_acetyl_peptide_norm_no_var <- liver_acetyl_peptide_norm %>% 
                                        dplyr::filter(zero_var) %>% 
                                        dplyr::select(-zero_var)

liver_acetyl_peptide_only <- liver_acetyl_peptide %>% 
  tidyr::separate_rows(`Master Protein Accessions`) %>% 
  tidyr::unite("unique_id", peptide_id:`Master Protein Accessions`) %>% 
  dplyr::filter(unique_id %in% liver_acetyl_peptide_norm_no_var$unique_id)

# combine data frames
# NOTE: treat separately for limma!
liver_acetyl_peptide_norm <- dplyr::bind_rows(liver_acetyl_peptide_norm_var,
                                              liver_acetyl_peptide_only)
```

### Heatmap

```{r message = FALSE, warning = FALSE}
heatmap_function(df = liver_acetyl_peptide_norm,
                 cols = c(3, 4, 2),
                 log_transform = FALSE,
                 cluster_number = 3,
                 distance_metric_row = "correlation",
                 distance_metric_col = "correlation",
                 agglomeration_method = "average",
                 cluster_cols = FALSE,
                 sample = c("10 mM", "1 mM", "0 mM"))
```

## Figures 2 and S2

Analysis of the cysteine acetylome following subcellular fractionation of mouse liver into supernatant, heavy, and light fractions. 

Volcano plots showing proteins and sites with the greatest differences in modification level between subcellular compartments.  

### Import data

```{r message = FALSE, warning = FALSE}
# import peptide-level data
subcell_peptide <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20230330EKK20230327_LiverCellFrac_PD20230403.xlsx")

# Add a unique identifier to each row
subcell_peptide$peptide_id <- paste0("analyte_", 1:nrow(subcell_peptide))

# import protein-level data
subcell_protein <- read_excel("Protein data/All Proteins _ all columns _ 20230330EKK20230327_LiverCellFrac_PD20230403.xlsx")

# In the case of peptide-level data, there are some peptides that match to MULTIPLE proteins
# Decision - treat these as DIFFERENT peptides

# select relevant columns 
subcell_peptide <- subcell_peptide %>% 
                            dplyr::select(peptide_id, `Master Protein Accessions`, starts_with("Abundances (Normalized)")) %>% 
                            tidyr::drop_na() %>% 
                   dplyr::rename_with(.fn = ~str_replace(., "Ctyo", "Supernatant"),
                     .cols = contains("Ctyo")) %>%
  dplyr::rename_with(.fn = ~str_replace(., "Mito", "Light"),
                     .cols = contains("Mito")) %>%
  dplyr::rename_with(.fn = ~str_replace(., "Nuc", "Heavy"),
                     .cols = contains("Nuc")) # update column names

# reshape data frame
subcell_peptide_long <- subcell_peptide %>%
  tidyr::separate_longer_delim(`Master Protein Accessions`,
                               delim = ";") %>%
  tidyr::pivot_longer(
    cols = starts_with("Abundance"),
    names_to = "sample_name",
    values_to = "peptide_value"
  ) %>%
  dplyr::rename(Accession = `Master Protein Accessions`) %>%
  dplyr::mutate(Accession = str_trim(Accession))
# select relevant columns
subcell_protein <- subcell_protein %>% 
                            dplyr::select(Accession, starts_with("Abundances (Normalized)")) %>% 
                            tidyr::drop_na() %>% 
                            dplyr::rename_with(.fn = ~str_replace(., "Ctyo", "Supernatant"),
                     .cols = contains("Ctyo")) %>%
  dplyr::rename_with(.fn = ~str_replace(., "Mito", "Light"),
                     .cols = contains("Mito")) %>%
  dplyr::rename_with(.fn = ~str_replace(., "Nuc", "Heavy"),
                     .cols = contains("Nuc")) # update column names

# reshape data frame
subcell_protein_long <- subcell_protein %>% 
  tidyr::pivot_longer(cols = starts_with("Abundance"),
                      names_to = "sample_name",
                      values_to = "protein_value") 

# join data frames, normalize peptide value, and reshape
subcell_peptide_norm <- subcell_peptide_long %>% 
  dplyr::inner_join(subcell_protein_long,
                    by = c("Accession", "sample_name")) %>% 
  dplyr::mutate(normalized_value = peptide_value/protein_value,
                   .keep = "unused") %>% 
  tidyr::pivot_wider(
    names_from = sample_name,
    values_from = normalized_value
  ) %>% 
  tidyr::unite("unique_id", peptide_id:Accession) # this way every entry will be unique

# Note that these are a few analytes that have 1s across all samples, indicating that in these cases the peptide and protein abundance are IDENTICAL
 
# create a column called `zero_var` that identifies those rows with 0 variance i.e. all 1s
subcell_peptide_norm <- subcell_peptide_norm %>% 
                                  dplyr::mutate(zero_var = if_all(starts_with("Abundances"), ~ . == 1))

subcell_peptide_norm_var <- subcell_peptide_norm %>% 
                                    dplyr::filter(!zero_var) %>% 
                                    dplyr::select(-zero_var)
subcell_peptide_norm_no_var <- subcell_peptide_norm %>% 
                                        dplyr::filter(zero_var) %>% 
                                        dplyr::select(-zero_var)

subcell_peptide_only <- subcell_peptide %>% 
  tidyr::separate_rows(`Master Protein Accessions`, sep = ";") %>% 
  tidyr::unite("unique_id", peptide_id:`Master Protein Accessions`) %>% 
  dplyr::filter(unique_id %in% subcell_peptide_norm_no_var$unique_id)

# combine data frames
# NOTE: treat separately for limma!
subcell_peptide_norm <- dplyr::bind_rows(subcell_peptide_norm_var,
                                               subcell_peptide_only)
```

### Heatmap

```{r message = FALSE, warning = FALSE}
heatmap_function(df = subcell_peptide_norm,
                 cols = c(2:10),
                 log_transform = FALSE,
                 cluster_number = 6, 
                 distance_metric_row = "correlation", 
                 distance_metric_col = "correlation", 
                 agglomeration_method = "average",
                 sample = c(rep("Supernatant", 3), rep("Light", 3), rep("Heavy", 3)))
```

### PCA 

```{r message = FALSE, warning = FALSE}
# on peptide data
subcell_peptide_pca_norm <- subcell_peptide_norm[2:10] %>% 
                                  tidyr::drop_na() %>% 
                                  as.matrix()
subcell_peptide_pca_norm <- t(subcell_peptide_pca_norm)
rownames(subcell_peptide_pca_norm) <- c(rep("Supernatant", 3), rep("Light", 3), rep("Heavy", 3))
subcell_peptide_res.pca_norm <- PCA(subcell_peptide_pca_norm,  graph = FALSE)

# Visualize eigenvalues/variances
fviz_screeplot(subcell_peptide_res.pca_norm, addlabels = TRUE, ylim = c(0, 50))

fviz_pca_ind(subcell_peptide_res.pca_norm, 
             habillage = as.factor(rownames(subcell_peptide_pca_norm)),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )

# on protein data
subcell_protein_pca <- subcell_protein[2:10] %>% 
                                  tidyr::drop_na() %>% 
                                  as.matrix()
subcell_protein_pca <- t(subcell_protein_pca)
rownames(subcell_protein_pca) <- c(rep("Supernatant", 3), rep("Light", 3), rep("Heavy", 3))
subcell_protein_res.pca_norm <- PCA(subcell_protein_pca,  graph = FALSE)

# Visualize eigenvalues/variances
fviz_screeplot(subcell_protein_res.pca_norm, addlabels = TRUE, ylim = c(0, 50))

fviz_pca_ind(subcell_protein_res.pca_norm, 
             habillage = as.factor(rownames(subcell_protein_pca)),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )
```

### Differential Abundance Analysis

```{r message = FALSE, warning = FALSE}
# first on set normalized to protein levels
# prepare matrix for limma and log2 transform
subcell_peptide_norm_var_mat <- as.matrix(subcell_peptide_norm_var[ , -c(1, 11, 12)])
subcell_peptide_norm_var_mat <- apply(subcell_peptide_norm_var_mat, 2, log2)
rownames(subcell_peptide_norm_var_mat) <- subcell_peptide_norm_var$unique_id

# create design matrix
subcell_group <- c(rep("Supernatant", 3), rep("Light", 3), rep("Heavy", 3))
subcell_group <- factor(subcell_group,
                                levels = c("Supernatant", "Light", "Heavy"))
subcell_design <- model.matrix(~0 + subcell_group)
colnames(subcell_design) <- c("Supernatant_mean", "Light_mean", "Heavy_mean")
cell_identity <- rep(c(1, 2, 3), 3)

cor_var <- duplicateCorrelation(subcell_peptide_norm_var_mat, 
                            subcell_design, 
                            block = cell_identity)
# cor_var$consensus.correlation
#-0.01765019

# fit model
subcell_var_fit <- limma::lmFit(subcell_peptide_norm_var_mat, 
                                subcell_design,
                                block = cell_identity,
                                correlation=cor_var$consensus.correlation)

# fit contrasts
subcell.cont <- makeContrasts(
  Light_v_Supernatant = Light_mean - Supernatant_mean,
  Heavy_v_Supernatant = Heavy_mean - Supernatant_mean,
  Heavy_v_Light = Heavy_mean - Light_mean,
  levels = subcell_design
)
subcell_var_fit_cont <- contrasts.fit(subcell_var_fit, subcell.cont)

subcell_var_efit <- limma::eBayes(subcell_var_fit_cont)

# now on "zero var" set
# prepare matrix for limma and log2 transform
subcell_peptide_only_mat <- as.matrix(subcell_peptide_only[ , -c(1, 11, 12)])
subcell_peptide_only_mat <- apply(subcell_peptide_only_mat, 2, log2)
rownames(subcell_peptide_only_mat) <- subcell_peptide_only$unique_id

cor_novar <- duplicateCorrelation(subcell_peptide_only_mat, 
                                  subcell_design, 
                                  block = cell_identity)
# cor_novar$consensus.correlation
#-0.09322086

# fit model
subcell_no_var_fit <- limma::lmFit(subcell_peptide_only_mat, 
                                   subcell_design,
                                   block = cell_identity,
                                   correlation = cor_novar$consensus.correlation)

# fit contrasts
subcell_no_var_fit_cont <- contrasts.fit(subcell_no_var_fit, subcell.cont)

subcell_no_var_efit <- limma::eBayes(subcell_no_var_fit_cont)

# bind results
# for Light vs Supernatant
light_vs_supernatant_results <- dplyr::bind_rows(topTable(subcell_var_efit, 
         coef = "Light_v_Supernatant",
         number = nrow(subcell_peptide_norm_var_mat)),
         topTable(subcell_no_var_efit, 
         coef = "Light_v_Supernatant",
         number = nrow(subcell_peptide_only_mat))) %>% 
            data.frame() %>% 
            tibble::rownames_to_column("unique_id") %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))

# for Heavy vs Supernatant
heavy_vs_supernatant_results <- dplyr::bind_rows(topTable(subcell_var_efit, 
         coef = "Heavy_v_Supernatant",
         number = nrow(subcell_peptide_norm_var_mat)),
         topTable(subcell_no_var_efit, 
         coef = "Heavy_v_Supernatant",
         number = nrow(subcell_peptide_only_mat))) %>% 
            data.frame() %>% 
            tibble::rownames_to_column("unique_id") %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))

# for Heavy vs Light
heavy_vs_light_results <- dplyr::bind_rows(topTable(subcell_var_efit, 
         coef = "Heavy_v_Light",
         number = nrow(subcell_peptide_norm_var_mat)),
         topTable(subcell_no_var_efit, 
         coef = "Heavy_v_Light",
         number = nrow(subcell_peptide_only_mat))) %>% 
            data.frame() %>% 
            tibble::rownames_to_column("unique_id") %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
         dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))

# Note - the adjusted p-values computed here are incorrect because of the way you initially split the data (and ran limma separately)
# replace these values with BH-corrected p-values computed using ALL rows
light_vs_supernatant_results$adj.P.Val <- p.adjust(light_vs_supernatant_results$P.Value,
                                              method = "BH")
heavy_vs_supernatant_results$adj.P.Val <- p.adjust(heavy_vs_supernatant_results$P.Value,
                                              method = "BH")
heavy_vs_light_results$adj.P.Val <- p.adjust(heavy_vs_light_results$P.Value,
                                              method = "BH")

# convert to proper name format
# e.g. MDH1-C123
protein_site <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20230330EKK20230327_LiverCellFrac_PD20230403.xlsx")
  
protein_site$peptide_id <- paste0("analyte_", 1:nrow(protein_site))

protein_site <- protein_site %>% 
  tidyr::separate_longer_delim(`Master Protein Accessions`,
                               delim = ";") %>%
  tidyr::separate_longer_delim(`Modifications in Master Proteins (all Sites)`,
                               delim = "];") %>%
  dplyr::rename(Accession = `Master Protein Accessions`) %>%
  dplyr::mutate(Accession = str_trim(Accession)) %>%
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = str_trim(`Modifications in Master Proteins (all Sites)`,
                                     side = "left")) %>%
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = paste0(`Modifications in Master Proteins (all Sites)`, "]")) %>% 
  # adding closing square brackets to those residues that are missing it. 
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = str_replace(`Modifications in Master Proteins (all Sites)`, "\\]\\]", "\\]")) %>% 
  # now everything looks good!
  dplyr::mutate(protein_residue = word(`Modifications in Master Proteins (all Sites)`)) %>% 
  dplyr::filter(Accession == protein_residue) %>% 
  dplyr::select(peptide_id, Accession, `Modifications in Master Proteins (all Sites)`) %>% 
  dplyr::rename(Modification = `Modifications in Master Proteins (all Sites)`) %>% 
  dplyr::mutate(Modification = str_extract(Modification, "(\\[).+(\\])")) %>%  # pull out anything between square brackets
  dplyr::mutate(Modification = str_remove_all(Modification, "(?<=\\]).+")) %>% # only keep things inside square brackets
  tidyr::unite("unique_id", peptide_id:Accession, sep = "-", remove = FALSE) %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))

protein_site <- biomaRt::select(org.Mm.eg.db, protein_site$Accession, "SYMBOL", "UNIPROT") %>% 
  tibble::tibble() %>% 
  dplyr::inner_join(protein_site, by = c("UNIPROT" = "Accession")) %>% 
  dplyr::mutate(SYMBOL = str_to_upper(SYMBOL)) %>% 
  tidyr::unite("protein_mod", c(SYMBOL, Modification), sep = "-") %>% 
  dplyr::mutate(protein_mod = if_else(str_detect(protein_mod, "^NA"),
                                      NA,
                                      protein_mod)) %>% 
  dplyr::select(unique_id, protein_mod)

# join
# for Light vs Supernatant
light_vs_supernatant_results_join <- light_vs_supernatant_results %>% 
  dplyr::inner_join(protein_site, by = "unique_id") %>% 
  dplyr::distinct()
# for Heavy vs Supernatant
heavy_vs_supernatant_results_join <- heavy_vs_supernatant_results %>% 
  dplyr::inner_join(protein_site, by = "unique_id") %>% 
  dplyr::distinct()
# for Heavy vs Light
heavy_vs_light_results_join <- heavy_vs_light_results %>% 
  dplyr::inner_join(protein_site, by = "unique_id") %>% 
  dplyr::distinct()

# plots
light_vs_supernatant_results_join %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-\\[.+")) %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-NA")) %>% 
  ggplot(aes(logFC, -log10(adj.P.Val))) +
  geom_point(col = "grey") +
  geom_vline(xintercept = -2, linetype = 2) +
  geom_vline(xintercept = 2, linetype = 2) +
  geom_hline(yintercept = -log10(0.1), linetype = 3) +
  geom_vline(xintercept = 0, linetype = 1) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_point(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), col = "red") +
  geom_point(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), col = "blue") +
  geom_text_repel(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = 1) +
  geom_text_repel(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = -0.1) +
  labs(title = "Light vs Supernatant") +
  theme_bw() +
  scale_x_continuous(limits = c(-10, 10)) +
  scale_y_continuous(limits = c(0, 4)) 

heavy_vs_supernatant_results_join %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-\\[.+")) %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-NA")) %>% 
  ggplot(aes(logFC, -log10(adj.P.Val))) +
  geom_point(col = "grey") +
  geom_vline(xintercept = -2, linetype = 2) +
  geom_vline(xintercept = 2, linetype = 2) +
  geom_hline(yintercept = -log10(0.1), linetype = 3) +
  geom_vline(xintercept = 0, linetype = 1) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_point(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), col = "red") +
  geom_point(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), col = "blue") +
  geom_text_repel(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = 1) +
  geom_text_repel(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = -0.1) +
  labs(title = "Heavy vs Supernatant") +
  theme_bw() +
  scale_x_continuous(limits = c(-10, 10)) +
  scale_y_continuous(limits = c(0, 4)) 

heavy_vs_light_results_join %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-\\[.+")) %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-NA")) %>% 
  ggplot(aes(logFC, -log10(adj.P.Val))) +
  geom_point(col = "grey") +
  geom_vline(xintercept = -2, linetype = 2) +
  geom_vline(xintercept = 2, linetype = 2) +
  geom_hline(yintercept = -log10(0.1), linetype = 3) +
  geom_vline(xintercept = 0, linetype = 1) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_point(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), col = "red") +
  geom_point(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), col = "blue") +
  geom_text_repel(data = . %>% filter(logFC > 2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = 1) +
  geom_text_repel(data = . %>% filter(logFC < -2 & adj.P.Val < 0.1), aes(label = protein_mod),
            vjust = -1, hjust = -0.1) +
  labs(title = "Heavy vs Light") +
  theme_bw() +
  scale_x_continuous(limits = c(-10, 10)) +
  scale_y_continuous(limits = c(0, 4)) 
```

### Venn diagram

```{r message = FALSE, warning = FALSE}
# import peptide-level data
subcell_peptide <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20230330EKK20230327_LiverCellFrac_PD20230403.xlsx")

# Add a unique identifier to each row
subcell_peptide$peptide_id <- paste0("analyte_", 1:nrow(subcell_peptide))

subcel_venn <- subcell_peptide %>% 
  dplyr::select(peptide_id, contains("Found in Sample Group")) %>% 
  dplyr::select(!contains("Total"))

light_group <- subcel_venn[subcel_venn[[3]] != "Not Found", ]$peptide_id
supernatant_group <- subcel_venn[subcel_venn[[2]] != "Not Found", ]$peptide_id
heavy_group <- subcel_venn[subcel_venn[[4]] != "Not Found", ]$peptide_id
  
# Venn diagram
venn.diagram(
  x = list(light_group, supernatant_group, heavy_group),
  category.names = c("Light" , "Supernatant" , "Heavy"),
  filename = 'results/subcell_frac_venn_diagram_061825.png',
  output = TRUE
)

# analytes found in all three
common_peptides <- intersect(intersect(light_group, supernatant_group), heavy_group)

# pull out these sequences
common_peptide_df <- subcell_peptide %>% 
  dplyr::select(peptide_id, Sequence, `Master Protein Accessions`) %>% 
  dplyr::filter(peptide_id %in% common_peptides)
```

## Figure 3

Multi-tissue survey of the murine cysteine acetylome. 

```{r message = FALSE, warning = FALSE}
# import peptide-level data and replace NAs with mean of remaining samples in name group
# remove any rows that still have NAs
# acetyl only
bat_acetyl <- read_excel("MultiTissueSurvey/BAT _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_BAT.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(bat_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(bat_acetyl)
# 1

brain_acetyl <- read_excel("MultiTissueSurvey/Brain _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Brain.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(brain_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(brain_acetyl)
# 1

heart_acetyl <- read_excel("MultiTissueSurvey/Heart _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Heart.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(heart_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(heart_acetyl)
# 1

kidney_acetyl <- read_excel("MultiTissueSurvey/Kidney _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Kidney.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(kidney_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(kidney_acetyl)
# 1

liver_acetyl <- read_excel("MultiTissueSurvey/Liver _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Liver.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(liver_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(liver_acetyl)
# 1

lung_acetyl <- read_excel("MultiTissueSurvey/Lung _ All Acetyl_Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Lung.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(lung_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(lung_acetyl)
# 1

pancreas_acetyl <- read_excel("MultiTissueSurvey/Pancreas _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen__gBeW.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(pancreas_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(pancreas_acetyl)
# 1

skm_acetyl <- read_excel("MultiTissueSurvey/Skm _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Skm.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(skm_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(skm_acetyl)
# 1

spleen_acetyl <- read_excel("MultiTissueSurvey/Spleen _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Spleen.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(spleen_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(spleen_acetyl)
# 1

total_acetyl <- read_excel("MultiTissueSurvey/Total _ All Acetyl-Cys Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Total.xlsx") %>% 
              dplyr::select(`Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
              dplyr::select(-row_means) %>% 
              tidyr::drop_na() %>% # drop any rows that only have NAs across all samples
              dplyr::group_by(`Modifications in Master Proteins (all Sites)`) %>% 
              dplyr::summarise(across(starts_with("Abundance"), sum)) %>% 
              dplyr::ungroup() # because some of these data frames have duplicates in these columns

# check uniqueness of peptides
# length(unique(total_acetyl$`Modifications in Master Proteins (all Sites)`))/nrow(total_acetyl)
# 1

# normalize to total peptide sums
bat_peptide_colsums <- read_excel("MultiTissueSurvey/BAT _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_BAT.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
bat_peptide_colsums_df <- data.frame(as.list(bat_peptide_colsums))

brain_peptide_colsums <- read_excel("MultiTissueSurvey/Brain _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Brain.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
brain_peptide_colsums_df <- data.frame(as.list(brain_peptide_colsums))

heart_peptide_colsums <- read_excel("MultiTissueSurvey/Heart _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Heart.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
heart_peptide_colsums_df <- data.frame(as.list(heart_peptide_colsums))

kidney_peptide_colsums <- read_excel("MultiTissueSurvey/Kidney _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Kidney.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE) # NOTE - no Abundance columns!
kidney_peptide_colsums_df <- data.frame(as.list(kidney_peptide_colsums))

liver_peptide_colsums <- read_excel("MultiTissueSurvey/Liver _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Liver.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
liver_peptide_colsums_df <- data.frame(as.list(liver_peptide_colsums))

lung_peptide_colsums <- read_excel("MultiTissueSurvey/Lung _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Lung.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
lung_peptide_colsums_df <- data.frame(as.list(lung_peptide_colsums))

pancreas_peptide_colsums <- read_excel("MultiTissueSurvey/Pancreas _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Pancreas.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
pancreas_peptide_colsums_df <- data.frame(as.list(pancreas_peptide_colsums))

skm_peptide_colsums <- read_excel("MultiTissueSurvey/Skm _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Skm.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
skm_peptide_colsums_df <- data.frame(as.list(skm_peptide_colsums))

spleen_peptide_colsums <- read_excel("MultiTissueSurvey/Spleen _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Spleen.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
spleen_peptide_colsums_df <- data.frame(as.list(spleen_peptide_colsums))

total_peptide_colsums <- read_excel("MultiTissueSurvey/Total _ All Peptide groups _ all columns _ 2024-01-04_AcetylCysTissueScreen_Total.xlsx") %>% 
                       dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)
total_peptide_colsums_df <- data.frame(as.list(total_peptide_colsums))

# Divide "Abundance:" columns by these colsums
bat_acetyl <- bat_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / bat_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

brain_acetyl <- brain_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / brain_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

heart_acetyl <- heart_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / heart_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

kidney_acetyl <- kidney_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / kidney_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>%
  dplyr::ungroup()

liver_acetyl <- liver_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / liver_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

lung_acetyl <- lung_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / lung_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

pancreas_acetyl <- pancreas_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / pancreas_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

skm_acetyl <- skm_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / skm_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

spleen_acetyl <- spleen_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / spleen_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

total_acetyl <- total_acetyl %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / total_peptide_colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

# # perform mean imputation for protein abundances
# bat_protein <- read_excel("MultiTissueSurvey/BAT _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_BAT.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# brain_protein <- read_excel("MultiTissueSurvey/Brain _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Brain.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# heart_protein <- read_excel("MultiTissueSurvey/Heart _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Heart.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# # kidney_protein <- read_excel("MultiTissueSurvey/Kidney _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Kidney.xlsx") %>% 
# #                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
# #                    dplyr::rowwise() %>% 
# #                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
# #                    dplyr::ungroup() %>% 
# #                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
# #                    dplyr::select(-row_means) %>% 
# #                    tidyr::drop_na()
# 
# liver_protein <- read_excel("MultiTissueSurvey/Liver _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Liver.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# lung_protein <- read_excel("MultiTissueSurvey/Lung _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Lung.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# pancreas_protein <- read_excel("MultiTissueSurvey/Pancreas _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Pancreas.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# skm_protein <- read_excel("MultiTissueSurvey/Skm _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Skm.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# spleen_protein <- read_excel("MultiTissueSurvey/Spleen _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Spleen.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# total_protein <- read_excel("MultiTissueSurvey/Total _ All Proteins _ all columns _ 2024-01-04_AcetylCysTissueScreen_Total.xlsx") %>% 
#                    dplyr::select(Accession, starts_with("Abundance:")) %>% 
#                    dplyr::rowwise() %>% 
#                    dplyr::mutate(row_means = mean(c_across(starts_with("Abundance:")), na.rm = TRUE)) %>% 
#                    dplyr::ungroup() %>% 
#                    dplyr::mutate(across(starts_with("Abundance:"), ~ coalesce(., row_means))) %>% 
#                    dplyr::select(-row_means) %>% 
#                    tidyr::drop_na()
# 
# # normalize peptide values to protein values
# # reshape peptide data frame
# bat_acetyl_long <- bat_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# brain_acetyl_long <- brain_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# heart_acetyl_long <- heart_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# # kidney_acetyl_long <- kidney_acetyl %>%
# #   tidyr::separate_longer_delim(`Master Protein Accessions`,
# #                                delim = ";") %>%
# #   tidyr::pivot_longer(
# #     cols = starts_with("Abundance"),
# #     names_to = "sample_name",
# #     values_to = "peptide_value"
# #   ) %>%
# #   dplyr::rename(Accession = `Master Protein Accessions`) %>%
# #   dplyr::mutate(Accession = str_trim(Accession))
# 
# liver_acetyl_long <- liver_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# lung_acetyl_long <- lung_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# pancreas_acetyl_long <- pancreas_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# skm_acetyl_long <- skm_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# spleen_acetyl_long <- spleen_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# total_acetyl_long <- total_acetyl %>%
#   tidyr::separate_longer_delim(`Master Protein Accessions`,
#                                delim = ";") %>%
#   tidyr::pivot_longer(
#     cols = starts_with("Abundance"),
#     names_to = "sample_name",
#     values_to = "peptide_value"
#   ) %>%
#   dplyr::rename(Accession = `Master Protein Accessions`) %>%
#   dplyr::mutate(Accession = str_trim(Accession))
# 
# # reshape protein data frame
# bat_protein_long <- bat_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# brain_protein_long <- brain_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# heart_protein_long <- heart_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# # kidney_protein_long <- kidney_protein %>% 
# #   tidyr::pivot_longer(cols = starts_with("Abundance"),
# #                       names_to = "sample_name",
# #                       values_to = "protein_value")
# 
# liver_protein_long <- liver_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# lung_protein_long <- lung_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# pancreas_protein_long <- pancreas_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# skm_protein_long <- skm_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# spleen_protein_long <- spleen_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# total_protein_long <- total_protein %>% 
#   tidyr::pivot_longer(cols = starts_with("Abundance"),
#                       names_to = "sample_name",
#                       values_to = "protein_value")
# 
# # join data frames, normalize peptide value, and reshape
# bat_acetyl_norm <- bat_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# brain_acetyl_norm <- brain_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# heart_acetyl_norm <- heart_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# # kidney_peptide_norm <- kidney_peptide_long %>% 
# #   dplyr::inner_join(bat_protein_long,
# #                     by = c("Accession", "sample_name")) %>% 
# #   dplyr::mutate(normalized_value = peptide_value/protein_value,
# #                    .keep = "unused") %>%
# #   tidyr::pivot_wider(
# #     names_from = sample_name,
# #     values_from = normalized_value
# #   )
# 
# liver_acetyl_norm <- liver_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# lung_acetyl_norm <- lung_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# pancreas_acetyl_norm <- pancreas_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# skm_acetyl_norm <- skm_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# spleen_acetyl_norm <- spleen_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )
# 
# total_acetyl_norm <- total_acetyl_long %>% 
#   dplyr::inner_join(bat_protein_long,
#                     by = c("Accession", "sample_name")) %>% 
#   dplyr::mutate(normalized_value = peptide_value/protein_value,
#                    .keep = "unused") %>%
#   tidyr::pivot_wider(
#     names_from = sample_name,
#     values_from = normalized_value
#   )

# combine all data frames
# left_joins
combo_acetyl <- bat_acetyl %>% 
  dplyr::full_join(brain_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>% 
  dplyr::full_join(heart_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>% 
  dplyr::full_join(kidney_acetyl, by = "Modifications in Master Proteins (all Sites)", 
                   keep = FALSE) %>% 
  dplyr::full_join(liver_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>% 
  dplyr::full_join(lung_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>%
  dplyr::full_join(pancreas_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>%
  dplyr::full_join(skm_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>%
  dplyr::full_join(spleen_acetyl, by = "Modifications in Master Proteins (all Sites)",
                   keep = FALSE) %>%
  dplyr::select(!starts_with("Master"))

# replace all NAs with 0
combo_acetyl[is.na(combo_acetyl)] <- 0
```

### Heatmap
```{r warning = FALSE, message = FALSE}
# create column annotations
my_sample_col <- data.frame(
    sample = factor(c(rep("Brown Adipose Tissue (BAT)", 3), rep("Brain", 3), rep("Heart", 3), rep("Kidney", 3), 
               rep("Liver", 3), rep("Lung", 3), rep("Pancreas", 3), rep("Skeletal Muscle (Skm)", 3), 
               rep("Spleen", 3)),
               levels = c("Brown Adipose Tissue (BAT)", "Brain",
                          "Heart", "Kidney", "Liver", "Lung", "Pancreas", "Skeletal Muscle (Skm)", "Spleen"))
  )

rownames(my_sample_col) <- colnames(combo_acetyl[ , -1])

pheatmap(mat = as.matrix(combo_acetyl[ , -1]),
         scale = "row",
         distance_metric_row = "correlation", 
         cluster_cols = FALSE,  
         agglomeration_method = "average",
         annotation_col = my_sample_col,
         show_colnames = FALSE,
         main = "Heatmap of normalized acetyl peptide abundance across a range of tissues")
```

### PCA 
```{r warning = FALSE, message = FALSE}
# Add 'Total' samples
combo_acetyl_total <- combo_acetyl %>% 
                          dplyr::full_join(total_acetyl, by = "Modifications in Master Proteins (all Sites)",
                          keep = FALSE)

# replace all NAs with 0
combo_acetyl_total[is.na(combo_acetyl_total)] <- 0

combo_acetyl_total_pca <- combo_acetyl_total[-1] %>% 
                      tidyr::drop_na() %>% 
                      as.matrix() 
combo_acetyl_total_pca <- t(combo_acetyl_total_pca)
rownames(combo_acetyl_total_pca) <- c(rep("BAT", 3), rep("Brain", 3), rep("Heart", 3), rep("Kidney", 3), 
               rep("Liver", 3), rep("Lung", 3), rep("Pancreas", 3), rep("Skm", 3), 
               rep("Spleen", 3),
               rep("Total", 5))
combo_acetyl_total_res.pca <- PCA(combo_acetyl_total_pca,  graph = FALSE)

# Visualize eigenvalues/variances
fviz_screeplot(combo_acetyl_total_res.pca, addlabels = TRUE, ylim = c(0, 20))

fviz_pca_ind(combo_acetyl_total_res.pca, 
             habillage = as.factor(rownames(combo_acetyl_total_pca)),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )

# Reviewer comment - Sample labels should be removed from the PCA plot (FIG3C) because the key clearly defines the shapes and the labels make the plot difficult to interpret. Also, please consider zooming in and blowing up the part of the plot where the majority of the samples cluster so that the audience can effectively evaluate the results.

# choice 1 - do everything in one plot
# main plot
pca_p1 <- fviz_pca_ind(combo_acetyl_total_res.pca, 
             habillage = as.factor(rownames(combo_acetyl_total_pca)),
             geom = "point"
             ) +
          theme(
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank()
          ) +
          labs(title = "")

# subplot
pca_p2 <- fviz_pca_ind(combo_acetyl_total_res.pca, 
             habillage = as.factor(rownames(combo_acetyl_total_pca)),
             geom = "point"
             ) +
  xlim(-4, 3) +
  ylim(-5, 3) +
  labs(x = "",
       y = "",
       title = "") +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) 

pca_p1 + 
  annotation_custom(ggplotGrob(pca_p2), xmin = 1, xmax = 14, ymin = 0, ymax = 14) +
  geom_rect(aes(xmin = 2.1, xmax = 13.7, ymin = 1.2, ymax = 12.5), color='black', alpha=0,
            linetype = "dashed") +
  geom_rect(aes(xmin = -4.1, xmax = 3, ymin = -5, ymax = 1), color='black', alpha=0) +
  geom_path(aes(x,y,group=grp), 
            data=data.frame(x = c(-4.1, 2.1, 3, 13.7), y = c(1, 12.5, -5, 1.2),grp = c(1,1,2,2)),
            linetype='dashed')
```

## Figure 4

Cysteine acetylome in BAT from mice acclimated to 6C compared to mice housed at RT. 

### Import data

```{r message = FALSE, warning = FALSE}
# import peptide-level data
rt_vs_sixdeg_peptide <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20230212EKK20230126_BAT_4degVsRT_PD20230217.xlsx")

# Add a unique identifier to each row
rt_vs_sixdeg_peptide$peptide_id <- paste0("analyte_", 1:nrow(rt_vs_sixdeg_peptide))

# import protein-level data
rt_vs_sixdeg_protein <- read_excel("Protein data/All Proteins _ all columns _ 20230212EKK20230126_BAT_4degVsRT_PD20230217.xlsx")

# import FULL peptide-level data
rt_vs_sixdeg_peptide_all <- read_excel("Proteomics Data/All Peptide groups _ all columns _ 20230212EKK20230126_BAT_4degVsRT_PD20230217.xlsx")

# In the case of peptide-level data, there are some peptides that match to MULTIPLE proteins
# Decision - treat these as DIFFERENT peptides

# select relevant columns and perform mean imputation by group for peptide df
rt_vs_sixdeg_peptide <- rt_vs_sixdeg_peptide %>% 
              dplyr::select(peptide_id, `Master Protein Accessions`, `Modifications in Master Proteins (all Sites)`, starts_with("Abundance:")) %>% 
              dplyr::rowwise() %>% 
              dplyr::mutate(row_means_6deg = mean(c_across(contains("4deg")), na.rm = TRUE),
                            row_means_RT = mean(c_across(contains("RT")), na.rm = TRUE)) %>% 
              dplyr::ungroup() %>% 
              dplyr::mutate(across(contains("4deg"), ~ coalesce(., row_means_6deg)),
                            across(contains("RT"), ~ coalesce(., row_means_RT))) %>% 
              dplyr::select(-starts_with("row_means")) %>% 
              tidyr::drop_na()

# divide each column by the colsums of the FULL peptide dataframe
# compute columns sums for "Abundance:" columns
colsums <- rt_vs_sixdeg_peptide_all %>% 
  dplyr::select(starts_with("Abundance:")) %>% 
  colSums(., na.rm = TRUE)

colsums_df <- data.frame(as.list(colsums))

# Divide "Abundance:" columns by these colsums
rt_vs_sixdeg_peptide <- rt_vs_sixdeg_peptide %>%
  dplyr::rowwise() %>%
  dplyr::mutate(c_across(starts_with("Abundance:")) / colsums_df, .keep = "unused") %>%
  dplyr::rename_with(.cols = contains(".."), ~str_replace_all(., "\\.+", ": ")) %>% # we do this because the line above changes the sample names
  dplyr::rename_with(.cols = starts_with("Abundance"), ~str_replace(., "Sample:", "Sample,")) %>% 
  dplyr::ungroup()

# select relevant columns and perform mean imputation by group for protein df
rt_vs_sixdeg_protein <- rt_vs_sixdeg_protein %>% 
                            dplyr::select(Accession, starts_with("Abundance:")) %>% 
                            dplyr::rowwise() %>% 
                            dplyr::mutate(row_means_6deg = mean(c_across(contains("4deg")), na.rm = TRUE),
                                          row_means_RT = mean(c_across(contains("RT")), na.rm = TRUE)) %>% 
                            dplyr::ungroup() %>% 
                            dplyr::mutate(across(contains("4deg"), ~ coalesce(., row_means_6deg)),
                                          across(contains("RT"), ~ coalesce(., row_means_RT))) %>% 
                            dplyr::select(-starts_with("row_means")) %>% 
                            tidyr::drop_na()

# reshape peptide data frame
rt_vs_sixdeg_peptide_long <- rt_vs_sixdeg_peptide %>%
  tidyr::separate_longer_delim(`Master Protein Accessions`,
                               delim = ";") %>%
  tidyr::pivot_longer(
    cols = starts_with("Abundance"),
    names_to = "sample_name",
    values_to = "peptide_value"
  ) %>%
  dplyr::rename(Accession = `Master Protein Accessions`) %>%
  dplyr::mutate(Accession = str_trim(Accession))

# reshape protein data frame
rt_vs_sixdeg_protein_long <- rt_vs_sixdeg_protein %>% 
  tidyr::pivot_longer(cols = starts_with("Abundance"),
                      names_to = "sample_name",
                      values_to = "protein_value")

# join data frames, normalize peptide value, and reshape
rt_vs_sixdeg_peptide_norm <- rt_vs_sixdeg_peptide_long %>% 
  dplyr::inner_join(rt_vs_sixdeg_protein_long,
                    by = c("Accession", "sample_name")) %>% 
  dplyr::mutate(normalized_value = peptide_value/protein_value,
                   .keep = "unused") %>%
  tidyr::pivot_wider(
    names_from = sample_name,
    values_from = normalized_value
  ) %>% 
  tidyr::unite("unique_id", peptide_id:Accession) # this way every entry will be unique

# Note that these are a few analytes that have 1s across all samples, indicating that in these cases the peptide and protein abundance are IDENTICAL
 
# create a column called `zero_var` that identifies those rows with 0 variance i.e. all 1s
rt_vs_sixdeg_peptide_norm <- rt_vs_sixdeg_peptide_norm %>% 
                                  dplyr::mutate(zero_var = if_all(starts_with("Abundance:"), ~ . == 1))

rt_vs_sixdeg_peptide_norm_var <- rt_vs_sixdeg_peptide_norm %>% 
                                    dplyr::filter(!zero_var) %>% 
                                    dplyr::select(-zero_var) %>% 
                                    dplyr::select(-`Modifications in Master Proteins (all Sites)`)
rt_vs_sixdeg_peptide_norm_no_var <- rt_vs_sixdeg_peptide_norm %>% 
                                        dplyr::filter(zero_var) %>% 
                                        dplyr::select(-zero_var) %>% 
                                        dplyr::select(-`Modifications in Master Proteins (all Sites)`)# empty df

rt_vs_sixdeg_peptide_only <- rt_vs_sixdeg_peptide %>% 
  tidyr::separate_rows(`Master Protein Accessions`) %>% 
  tidyr::unite("unique_id", peptide_id:`Master Protein Accessions`) %>% 
  dplyr::filter(unique_id %in% rt_vs_sixdeg_peptide_norm_no_var$unique_id) %>% 
  dplyr::select(-`Modifications in Master Proteins (all Sites)`)

# combine data frames
# NOTE: treat separately for limma!
rt_vs_sixdeg_peptide_norm <- dplyr::bind_rows(rt_vs_sixdeg_peptide_norm_var,
                                                 rt_vs_sixdeg_peptide_only)
```

### Heatmap

```{r message = FALSE, warning = FALSE}
heatmap_function(df = rt_vs_sixdeg_peptide_norm,
                 cols = c(5,6,7,2,3,4),
                 log_transform = FALSE,
                 cluster_number = 6,
                 distance_metric_row = "correlation",
                 distance_metric_col = "correlation",
                 agglomeration_method = "average",
                 sample = factor(c(rep("RT", 3), rep("6deg", 3)),
                                 levels = c("RT", "6deg"))
                 )
```

### PCA 

```{r message = FALSE, warning = FALSE}
rt_vs_sixdeg_peptide_pca_norm <- rt_vs_sixdeg_peptide_norm[2:7] %>% 
                                  tidyr::drop_na() %>% 
                                  as.matrix() 
rt_vs_sixdeg_peptide_pca_norm <- t(rt_vs_sixdeg_peptide_pca_norm)
rownames(rt_vs_sixdeg_peptide_pca_norm) <- c(rep("6deg", 3), rep("RT", 3))
rt_vs_sixdeg_peptide_res.pca_norm <- PCA(rt_vs_sixdeg_peptide_pca_norm,  graph = FALSE)

# Visualize eigenvalues/variances
fviz_screeplot(rt_vs_sixdeg_peptide_res.pca_norm, addlabels = TRUE, ylim = c(0, 75))

fviz_pca_ind(rt_vs_sixdeg_peptide_res.pca_norm, 
             habillage = as.factor(rownames(rt_vs_sixdeg_peptide_pca_norm)),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )
```

### Differential Abundance Analysis
```{r message = FALSE, warning = FALSE}
# first on set normalized to protein levels
# prepare matrix for limma and log2 transform
rt_vs_sixdeg_peptide_norm_var_mat <- as.matrix(rt_vs_sixdeg_peptide_norm_var[ , -1])
rt_vs_sixdeg_peptide_norm_var_mat <- apply(rt_vs_sixdeg_peptide_norm_var_mat, 2, log2)
rownames(rt_vs_sixdeg_peptide_norm_var_mat) <- rt_vs_sixdeg_peptide_norm_var$unique_id

# create design matrix
rt_vs_sixdeg_group <- c(rep("6deg", 3), rep("RT", 3))
rt_vs_sixdeg_group <- factor(rt_vs_sixdeg_group,
                                levels = c("RT", "6deg"))
rt_vs_sixdeg_design <- model.matrix(~rt_vs_sixdeg_group)
colnames(rt_vs_sixdeg_design) <- c("RT_mean", "6deg_vs_RT")

# fit model
rt_vs_sixdeg_peptide_norm_var_mat_fit <- limma::lmFit(rt_vs_sixdeg_peptide_norm_var_mat, rt_vs_sixdeg_design)
rt_vs_sixdeg_peptide_norm_var_mat_efit <- limma::eBayes(rt_vs_sixdeg_peptide_norm_var_mat_fit)

# combine results
# note: here we're not combining anything because there are no "no var" analytes, unlike with the other rt_vs_6deg analysis
# however, the object names are kept the same for simplicity
rt_vs_sixdeg_combo <- topTable(rt_vs_sixdeg_peptide_norm_var_mat_efit, 
         coef = "6deg_vs_RT",
         number = nrow(rt_vs_sixdeg_peptide_norm_var_mat)) %>% 
  data.frame() %>% 
  tibble::rownames_to_column("unique_id") %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))
  
# convert to proper name format
# e.g. MDH1-C123
protein_site <- read_excel("Proteomics Data/All Acetyl-Cys Peptide groups _ all columns _ 20230212EKK20230126_BAT_4degVsRT_PD20230217.xlsx")
  
protein_site$peptide_id <- paste0("analyte_", 1:nrow(protein_site))

protein_site <- protein_site %>% 
  tidyr::separate_longer_delim(`Master Protein Accessions`,
                               delim = ";") %>%
  tidyr::separate_longer_delim(`Modifications in Master Proteins (all Sites)`,
                               delim = "];") %>%
  dplyr::rename(Accession = `Master Protein Accessions`) %>%
  dplyr::mutate(Accession = str_trim(Accession)) %>%
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = str_trim(`Modifications in Master Proteins (all Sites)`,
                                     side = "left")) %>%
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = paste0(`Modifications in Master Proteins (all Sites)`, "]")) %>% 
  # adding closing square bracked to those residues that are missing it. 
  dplyr::mutate(`Modifications in Master Proteins (all Sites)` = str_replace(`Modifications in Master Proteins (all Sites)`, "\\]\\]", "\\]")) %>% 
  # now everything looks good!
  dplyr::mutate(protein_residue = word(`Modifications in Master Proteins (all Sites)`)) %>% 
  dplyr::filter(Accession == protein_residue) %>% 
  dplyr::select(peptide_id, Accession, `Modifications in Master Proteins (all Sites)`) %>% 
  dplyr::rename(Modification = `Modifications in Master Proteins (all Sites)`) %>% 
  dplyr::mutate(Modification = str_extract(Modification, "(\\[).+(\\])")) %>%  # pull out anything between square bracket
  dplyr::mutate(Modification = str_remove_all(Modification, "(?<=\\]).+")) %>% # only keep things inside square brackets
  tidyr::unite("unique_id", peptide_id:Accession, sep = "-", remove = FALSE) %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_"))

protein_site <- biomaRt::select(org.Mm.eg.db, protein_site$Accession, "SYMBOL", "UNIPROT") %>% 
  tibble::tibble() %>% 
  dplyr::inner_join(protein_site, by = c("UNIPROT" = "Accession")) %>% 
  dplyr::mutate(SYMBOL = str_to_upper(SYMBOL)) %>% 
  tidyr::unite("protein_mod", c(SYMBOL, Modification), sep = "-") %>% 
  dplyr::mutate(protein_mod = if_else(str_detect(protein_mod, "^NA"),
                                      NA,
                                      protein_mod)) %>% 
  dplyr::select(unique_id, protein_mod)

# join
rt_vs_sixdeg_combo_join <- rt_vs_sixdeg_combo %>% 
  dplyr::inner_join(protein_site, by = "unique_id") %>% 
  dplyr::distinct()

rt_vs_sixdeg_combo_join %>% 
  dplyr::mutate(protein_mod = str_remove(protein_mod, "-\\[.+")) %>% 
  ggplot(aes(logFC, -log10(adj.P.Val))) +
  geom_point(col = "grey") +
  geom_vline(xintercept = -2, linetype = 2) +
  geom_vline(xintercept = 2, linetype = 2) +
  geom_hline(yintercept = -log10(0.05), linetype = 3) +
  geom_vline(xintercept = 0, linetype = 1) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_point(data = . %>% filter(logFC > 2 & adj.P.Val < 0.05), col = "red") +
  geom_point(data = . %>% filter(logFC < -2 & adj.P.Val < 0.05), col = "blue") +
  geom_text_repel(data = . %>% filter(logFC > 2 & adj.P.Val < 0.05), aes(label = protein_mod),
            vjust = -1, hjust = 1) +
  geom_text_repel(data = . %>% filter(logFC < -2 & adj.P.Val < 0.05), aes(label = protein_mod),
            vjust = -1, hjust = -0.1) +
  labs(title = "6deg vs RT") +
  theme_bw()

# visualize abundance of top analytes
rt_vs_sixdeg_peptide_norm %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)-(?=\\d)", "_")) %>% 
  dplyr::inner_join(protein_site, by = "unique_id") %>% 
  dplyr::select(-unique_id) %>%
  tidyr::pivot_longer(cols = -protein_mod,
                      names_to = "sample",
                      values_to = "abundance") %>% 
  dplyr::filter(protein_mod == "GAPDH-[K143; C150]") %>% 
  dplyr::mutate(treatment = if_else(str_detect(sample, "4deg"),
                                    "6deg",
                                    "RT"),
                treatment = factor(treatment),
                log_abundance = log2(abundance)) %>% 
  ggplot(aes(treatment, log_abundance, col = treatment)) +
  geom_point() +
  labs(title = "Analyte: GAPDH-[K143; C150]") +
  theme_bw()
```

### Pathway Analysis
```{r message = FALSE, warning = FALSE}
rt_vs_sixdeg_peptide_norm_fgsea <- topTable(rt_vs_sixdeg_peptide_norm_var_mat_efit, 
         coef = "6deg_vs_RT",
         number = nrow(rt_vs_sixdeg_peptide_norm_var_mat)) %>% 
  data.frame() %>% 
  tibble::rownames_to_column("unique_id") %>% 
  dplyr::select(unique_id, t) %>%
  dplyr::distinct() %>% 
  dplyr::mutate(unique_id = str_replace(unique_id, "(?<=\\d)_(?=\\w)", "-")) %>% 
  tidyr::separate_wider_delim(unique_id,
                              delim = "-",
                              names = c("analyte_id", "UNIPROT")) %>% 
  dplyr::select(UNIPROT, t) %>%
  dplyr::group_by(UNIPROT) %>% 
  dplyr::summarise(mean_t = mean(t))

# convert Uniprot IDs to Entrez IDs
# https://www.biostars.org/p/399332/

rt_vs_sixdeg_peptide_norm_entrez <- select(org.Mm.eg.db, rt_vs_sixdeg_peptide_norm_fgsea$UNIPROT, "ENTREZID", "UNIPROT")

# join
rt_vs_sixdeg_peptide_norm_entrez_ranks <- rt_vs_sixdeg_peptide_norm_fgsea %>% 
                          dplyr::inner_join(rt_vs_sixdeg_peptide_norm_entrez, by = "UNIPROT") %>% 
                          dplyr::select(ENTREZID, mean_t) %>% 
                          dplyr::arrange(desc(mean_t)) %>% 
                          tidyr::drop_na()

rt_vs_sixdeg_peptide_norm_entrez_ranks <- deframe(rt_vs_sixdeg_peptide_norm_entrez_ranks)

# gmt file obtained from - 
# https://data.broadinstitute.org/gsea-msigdb/msigdb/release/2023.1.Mm/

pathways.mh.all.v2023 <- gmtPathways("metadata/mh.all.v2023.1.Mm.entrez.gmt.txt")

set.seed(42)
fgseaRes_rt_vs_sixdeg_peptide_norm <- fgseaMultilevel(pathways = pathways.mh.all.v2023, 
                                            stats = rt_vs_sixdeg_peptide_norm_entrez_ranks)

fgseaResTidy_rt_vs_sixdeg_peptide_norm  <- fgseaRes_rt_vs_sixdeg_peptide_norm %>%
                                    tibble::as_tibble() %>%
                                    dplyr::arrange(padj)
```

### GSEA plot for `HALLMARK_ADIPOGENESIS`
```{r message = FALSE, warning = FALSE}
# gsea plot for Hallmark Adipogensis (the top pathway)
plotEnrichment(pathways.mh.all.v2023[["HALLMARK_ADIPOGENESIS"]],
               rt_vs_sixdeg_peptide_norm_entrez_ranks) +
  labs(title = "Adipogenesis")
```

### GSEA plot for `HALLMARK_FATTY_ACID_METABOLISM`
```{r message = FALSE, warning = FALSE}
plotEnrichment(pathways.mh.all.v2023[["HALLMARK_FATTY_ACID_METABOLISM"]],
               rt_vs_sixdeg_peptide_norm_entrez_ranks) +
  labs(title = "Fatty Acid Metabolism")
```

### GSEA plot for `HALLMARK_OXIDATIVE_PHOSPHORYLATION`
```{r message = FALSE, warning = FALSE}
plotEnrichment(pathways.mh.all.v2023[["HALLMARK_OXIDATIVE_PHOSPHORYLATION"]],
               rt_vs_sixdeg_peptide_norm_entrez_ranks) +
  labs(title = "Oxidative Phosphorylation")
```

### Pull out leading edge genes for `HALLMARK_ADIPOGENESIS`, `HALLMARK_FATTY_ACID_METABOLISM`, and `HALLMARK_OXIDATIVE_PHOSPHORYLATION`
```{r message = FALSE, warning = FALSE}
# `HALLMARK_ADIPOGENESIS`
fgseaResTidy_rt_vs_sixdeg_peptide_norm[fgseaResTidy_rt_vs_sixdeg_peptide_norm$pathway == "HALLMARK_ADIPOGENESIS", ]$leadingEdge
# "11429" "11770"
rt_vs_sixdeg_peptide_norm_entrez[rt_vs_sixdeg_peptide_norm_entrez$ENTREZID %in% 
                                    c("11429", "11770"), ]$UNIPROT
# "P04117" "Q99KI0"

# `HALLMARK_FATTY_ACID_METABOLISM`
fgseaResTidy_rt_vs_sixdeg_peptide_norm[fgseaResTidy_rt_vs_sixdeg_peptide_norm$pathway == "HALLMARK_FATTY_ACID_METABOLISM", ]$leadingEdge
# "11429" "16852" "72479"
rt_vs_sixdeg_peptide_norm_entrez[rt_vs_sixdeg_peptide_norm_entrez$ENTREZID %in% 
                                    c("11429", "16852", "72479"), ]$UNIPROT
# "P16045" "Q2TPA8" "Q99KI0"

# `HALLMARK_OXIDATIVE_PHOSPHORYLATION`
fgseaResTidy_rt_vs_sixdeg_peptide_norm[fgseaResTidy_rt_vs_sixdeg_peptide_norm$pathway == "HALLMARK_OXIDATIVE_PHOSPHORYLATION", ]$leadingEdge
# "11429" "97212"
rt_vs_sixdeg_peptide_norm_entrez[rt_vs_sixdeg_peptide_norm_entrez$ENTREZID %in% 
                                    c("11429", "97212"), ]$UNIPROT
# "Q8BMS1" "Q99KI0"
```